Efficiently stable presentations from error-correcting codes

### 2025-10-21 daj-revision2/referee-response.txt

We thank the reviewer for their thorough comments. Most of the changes have been made, although we decided to not follow the referee's suggestion regarding the use of Lemma 2.7, see the response to point 21 below.

Follows a redacted copy of the referee's comments. To follow the conversation it may be helpful to pull up the original referee comments pdf ; but we found it more helpful to provide responses inline, hence the txt format.

----------
[...]

### 2025-10-21 daj-revision2/efficient_stability.tex
again
final resubmitted version

\begin{lemma}\label{lem:com-code}
Let $R''_{jj'}$ be the commutation relation $[x_j,x_{j'}]=e$. Then
\[ \Z_2^k \,=\, \big\langle x_1,\ldots,x_n \,:\, x_j^2=e\,,\; R_i\,,\; R''_{jj'}\,,\quad \forall 1\leq i\leq m,\, 1\leq j< j' \leq n\big\rangle\;.\]
\end{lemma}

\begin{proof}
The group defined by the right-hand side is obviously abelian and a $2$-group, so it is of the form $\Z_2^{k'}$ for some $k'$. In fact, it is isomorphic to the quotient of $\Z_2^n$ by the subgroup generated by the $\prod_{1\leq j \leq n} x_j^{h_{ij}}$. So it is $\Z_2^k$ where $k=n-\dim\textrm{im}\ h = \dim\ker h=\dim \code$.  
\end{proof}

@@

Let $G_\bRM = G(h_\bRM)$ be the group that is presented from $h_\bRM$ (recall from \Cref{sec:pres-code} that codes $\code_\bRM=\ker h_\bRM$ give rise to group presentations through the general construction~\eqref{eq:def-gh-pres}). We do not know if $G_\bRM = \Z_2^k$, with $k=t(d+1)^m$. Instead we modify the presentation $G(h_\bRM)$ by adding pairwise commutation relations, as in~\eqref{eq:wtgh-intro}. Let \tnote{Removed $\{R^\sq_k\}$ below following reviewer suggestion}
\[ G(h_\bRM) \,=\, \big\langle x_1,\ldots,x_N \;:\;  \{R^\ld_k\}\,,\; \{R^\had_{k}\} \big\rangle\;,\]
where  $R^\ld_k$ ranges over all relations implied by the ``low-degree test'' in Figure~\ref{fig:bRM-tester} and $R^\had_{k}$ ranges over all the relations implied by the ``Hadamard test.'' For $k=(i,j)\in\{1,\ldots,N\}^2$ such that $i<j$ let $R^\com_k$ be the relation $[x_i,x_j]=e$, where $[a,b]=aba^{-1}b^{-1}$ is the group commutator.
Then we define 
\begin{equation}\label{eq:z2k-eff}
\tilde{G}:=\widetilde{G(h_\bRM) } \,=\,\big\langle x_1,\ldots,x_N \;:\;  \{R^\ld_k\}\,,\; \{R^\had_{k}\}\, , \; \{ R^\com_k\}\big\rangle\;.
\end{equation}
From Lemma~\ref{lem:com-code} it follows that $\tilde{G}=\Z_2^k$. Our main result is an efficient stability result for this presentation. To state this we need to introduce distributions $\mu_S$ and $\mu_R$ on the generators and relations of $\tilde{G}$. The distribution $\mu_R$ is obtained as follows. With probability $1/3$ each, a relation from $\{R^\ld_k\}$ or $\{R^\had_k\}$ is chosen uniformly at random. With probability $1/3$, a random commutation relation from $R^\com_k$ is chosen according to the uniform mixture of the following two distributions:
\begin{enumerate}
\item For the first distribution, we select $u\in \F_q^m$ uniformly at random, $j\in\{1,\ldots,m\}$ uniformly at random, and $i\neq i'\in\{0,\ldots,d\}$ uniformly at random. Then select $\alpha,\beta\in \F_2^t$ uniformly at random and check commutation between $x_{u+t_i e_j,\alpha}$ and $x_{u+t_{i'} e_j,\beta}$, where we interpret the subscripts  $u+t_i e_j,\alpha$ and $u+t_{i'} e_j,\beta$ as corresponding to some integer in $[N]= (\F_2^t)^m \times \F_2^t$ (as sets!).

@@

\begin{theorem}\label{thm:z2-stab}
Let $\mC$ be the class of all tracial von Neumann algebras. 
The presentation of  $\Z_2^k$ given in~\eqref{eq:z2k-eff}, together with the distributions $\mu_S$ and $\mu_R$ defined above, is $(\delta,\mC)$ stable with 
\begin{equation}\label{eq:th-min}
 \delta(\eps) = \min\Big\{ C (mdt)^{c_1} (\eps^{c_2}+(1/q)^{c_3})\ ,\ C'_{t,d,m}\ \eps\Big\}\;,
\end{equation}
where $C,c_1,c_2,c_3$ are universal positive constants and $C'_{t,d,m}$ depends on $t,d,m$ but not $\eps$.\footnote{We do not attempt to compute the dependence of $C'_{t,d,m}$ on the parameters because we are more interested in the regime of parameters where the first part of the min is relevant. Nevertheless, we include it so that $\delta(\eps)\to_{\eps\to 0} 0$, as required.}   
\end{theorem}

>>>>

\begin{lemma}\label{lem:com-code}
Let $R''_{jj'}$ be the commutation relation $[x_j,x_{j'}]=e$. Then $\Z_2^k$ is isomorphic to 
\[ \Z_2^k \,\simeq\, \big\langle x_1,\ldots,x_n \,:\, x_j^2=e\,,\; R_i\,,\; R''_{jj'}\,,\quad \forall 1\leq i\leq m,\, 1\leq j< j' \leq n\big\rangle\;.\]
\end{lemma}

\begin{proof}
The group defined by the right-hand side is obviously abelian and a $2$-group, so it is of the form $\Z_2^{k'}$ for some $k'$. In fact, it is isomorphic to the quotient of $\Z_2^n$ by the subgroup generated by the $\prod_{1\leq j \leq n} x_j^{h_{ij}}$. So it is isomorphic to $\Z_2^k$ where $k=n-\dim\textrm{im}\ h = \dim\ker h=\dim \code$.  
\end{proof}

@@

Let $G_\bRM = G(h_\bRM)$ be the group that is presented from $h_\bRM$ (recall from \Cref{sec:pres-code} that codes $\code_\bRM=\ker h_\bRM$ give rise to group presentations through the general construction~\eqref{eq:def-gh-pres}). We do not know if $G_\bRM = \Z_2^k$, with $k=t(d+1)^m$. Instead we modify the presentation $G(h_\bRM)$ by adding pairwise commutation relations, as in~\eqref{eq:wtgh-intro}. Let \tnote{Removed $\{R^\sq_k\}$ below following reviewer suggestion}
\[ G(h_\bRM) \,=\, \big\langle x_1,\ldots,x_N \;:\;  \{R^\ld_k\}\,,\; \{R^\had_{k}\} \big\rangle\;,\]
where  $R^\ld_k$ ranges over all relations implied by the ``low-degree test'' in Figure~\ref{fig:bRM-tester} and $R^\had_{k}$ ranges over all the relations implied by the ``Hadamard test.'' For $k=(i,j)\in\{1,\ldots,N\}^2$ such that $i<j$ let $R^\com_k$ be the relation $[x_i,x_j]=e$, where $[a,b]=aba^{-1}b^{-1}$ is the group commutator.
Then we define 
\begin{equation}\label{eq:z2k-eff}
\tilde{G}:=\widetilde{G(h_\bRM) } \,=\,\big\langle x_1,\ldots,x_N \;:\;  \{R^\ld_k\}\,,\; \{R^\had_{k}\}\, , \; \{ R^\com_k\}\big\rangle\;.
\end{equation}
From Lemma~\ref{lem:com-code} it follows that $\tilde{G}$ is isomorphic to $\Z_2^k$. Our main result is an efficient stability result for this presentation. To state this we need to introduce distributions $\mu_S$ and $\mu_R$ on the generators and relations of $\tilde{G}$. The distribution $\mu_R$ is obtained as follows. With probability $1/3$ each, a relation from $\{R^\ld_k\}$ or $\{R^\had_k\}$ is chosen uniformly at random. With probability $1/3$, a random commutation relation from $R^\com_k$ is chosen according to the uniform mixture of the following two distributions:
\begin{enumerate}
\item For the first distribution, we select $u\in \F_q^m$ uniformly at random, $j\in\{1,\ldots,m\}$ uniformly at random, and $i\neq i'\in\{0,\ldots,d\}$ uniformly at random. Then select $\alpha,\beta\in \F_2^t$ uniformly at random and check commutation between $x_{u+t_i e_j,\alpha}$ and $x_{u+t_{i'} e_j,\beta}$, where we interpret the subscripts  $u+t_i e_j,\alpha$ and $u+t_{i'} e_j,\beta$ as corresponding to some integer in $[N]= (\F_2^t)^m \times \F_2^t$ (as sets!).

@@

\begin{theorem}\label{thm:z2-stab}
Let $\mC$ be the class of all tracial von Neumann algebras. 
The presentation of  $\Z_2^k$ given in~\eqref{eq:z2k-eff}, together with the distributions $\mu_S$ and $\mu_R$ defined above, is $(\delta,\mC)$ stable with 
\begin{equation}\label{eq:th-min}
 \delta(\eps) = \min\Big\{ C (mdt)^{c_1} (\eps^{c_2}+(1/q)^{c_3})\ ,\ C'_{t,d,m}\ \eps\Big\}\;,
\end{equation}
where $C,c_1,c_2,c_3$ are universal positive constants and $C'_{t,d,m}$ depends on $t,d,m$ but not $\eps$.\footnote{We do not attempt to compute the dependence of $C'_{t,d,m}$ on $t,d,m$ because we are more interested in the regime of parameters where the first part of the min in~\eqref{eq:th-min} is relevant. Nevertheless, we include it so that $\delta(\eps)\to_{\eps\to 0} 0$ as required.}   
\end{theorem}
